---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Load package 

```{r}
library(osmextractr)
```

The packages is composed by 4 main functions: 

1. `osmext_match`: Match the input zone with one of the files stored by the OSM providers
2. `osmext_download`: Download the chosen file
3. `osmext_vectortranslate`: Convert the pbf format into gpkg
4. `osmext_read`: Read the gpkg file

# Test `osmext_match`

The simplest case: 

```{r}
osmext_match("Italy")
```

The input `place` can be also specified using an `sfc_POINT` object with arbitrary CRS as documented in the following example. If there are multiple matches, the function returns the smallest area (according to the `level` variable). I would ignore the CRS warning for the moment. 

```{r}
coords_milan = sf::st_point(c(1514924.21, 5034552.92))
st_sfc_milan = sf::st_sfc(coords_milan, crs = 3003)
osmext_match(st_sfc_milan)
```

<!-- The input `place` can be also specified using a numeric vector of coordinates. In that case the CRS is assumed to be 4326:  -->

```{r, eval = FALSE}
osmext_match("Italy", provider = "bbbike") #TODO
osmext_match(c(9, 45)) #TODO, crs = 4326 is implicit in this case

osmext_match("US")
osmext_match("US", match_by = "iso3166_1_alpha2") # matching by iso3166 is really powerful IMO but it should be documented

osmext_match("Korea") # Add one function to explore matches with grep
osmext_match("Russia") # Add one function to explore matches with grep
osmext_match("RU", match_by = "iso3166_1_alpha2")
osmext_match("Isle Wight", max_string_dist = 3)
```


```{r, eval = FALSE}
```

# Test download

```{r, eval = FALSE}
iow = osmext_match("Isle Wight", max_string_dist = 3)
osmext_download(
  file_url = iow$pbf_url, 
  file_size = iow$pbf_file_size
)
```
