---
title: "Add new OpenStreetMap providers"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Add new OpenStreetMap providers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(osmextractr)
```

As of summer 2020 there are 7 services providing bulk OSM datasets listed at [https://wiki.openstreetmap.org/wiki/Processed_data_providers](https://wiki.openstreetmap.org/wiki/Processed_data_providers).
This package is designed to make it easy to add new providers.
To add a new provider there are three main steps, creating the zones, adding the provider and documentation, outlined below.

# Adding a `provider_zones` object to the package

The first and hardest step is to create an R object analogous to the `test_zones` object shown below:

```{r}
names(test_zones)
test_zones[c(2, 6)]
```

The output shows the three most important column names in dataset:

1. The dataset name
2. The url endpoint where .pbf files associated with each zone can be downloaded
3. The geometry, representing the spatial extent of the dataset

The best way to start creating a new `_zones` object for a new provider is probably by looking at the code we wrote for the first and [test](https://github.com/ITSLeeds/osmextractr/commit/be3b48e7ed7ccd957e988bacad053161247b206d) providers in [`data-raw/geofabrik_zones.R`](https://github.com/ITSLeeds/osmextractr/blob/master/data-raw/geofabrik_zones.R)
The following commands will clone this repo and open the relevant file:

```bash
git clone git@github.com:itsleeds/osmextractr
rstudio osmextractr/osmextractr.Rproj
```

Then in RStudio:

```{r, eval=FALSE}
file.edit("data-raw/geofabrik_zones.R")
```

Create a new script to document the code that generates the new object, e.g. for `bbbike`:

```{r, eval=FALSE}
file.edit("data-raw/bbbike_zones.R")
```

After you have created the new provider `_zones` file it's time to add the provider to the package.

# Adding the new provider to the package

Once you have created your overview `_zones` file as outlined in the previous step, you need to modify the following files for the provider to be available for others:

- [data.R](https://github.com/ITSLeeds/osmextractr/blob/master/R/data.R), where you'll need to document the new dataset
- [globals.R](https://github.com/ITSLeeds/osmextractr/blob/master/R/globals.R), where you'll need to add the new object name
- [match.RR](https://github.com/ITSLeeds/osmextractr/blob/master/R/match.RR), where you'll need to add the new object name

# Documenting the provider

The final step is also the most fun: documenting and using the provider.
Add an example, mention it in the README and tell others about what this new provider can do!

# Conclusion

This vignette talks through the main steps needed to extend `osmextractr` by adding new OSM data providers.
To see the same information in code form, see the PR that implemented the `bbike` provider here: https://github.com/ITSLeeds/osmextractr/pull/57/files
