---
title: "Add new OpenStreetMap providers"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Add new OpenStreetMap providers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette aims to provide a simple guide on adding a new provider to `osmextract`. 
Let' start loading the package: 

```{r setup}
library(osmextract)
```

As of summer 2020, there are several services providing bulk OSM datasets listed  [here](https://wiki.openstreetmap.org/wiki/Processed_data_providers) and [here](https://wiki.openstreetmap.org/wiki/Planet.osm#Country_and_area_extracts). 
At the moment, we support the following providers: 

```{r}
oe_providers()
```

This package is designed to make it easy to add new providers. 
There are three main steps to add a new provider: creating the zones, adding the provider and documenting it. 
They are outlined below. 

# Adding a `provider_zones` object to the package

The first and hardest step is to create an `sf` object analogous to the `test_zones` object shown below:

```{r}
names(test_zones)
str(test_zones[, c(2, 6, 7)])
```

The output shows the three most important column names:

1. The zone `name` (that is used for matching the input `place`, see `oe_match()`); 
1. The URL endpoint where `.pbf` files associated with each zone can be downloaded; 
1. The geometry, representing the spatial extent of the dataset.

The object must also include the fields `level` and `id`, which are used, respectively, for spatial matching and updating. 
See `oe_match()` and `oe_update()`. 

The best way to start creating a new `_zones` object for a new provider is probably by looking at the code we wrote for the first supported provider in [`data-raw/geofabrik_zones.R`](https://github.com/ITSLeeds/osmextract/blob/master/data-raw/geofabrik_zones.R). 
The following commands will clone this repo and open the relevant file:

```bash
git clone git@github.com:itsleeds/osmextract
rstudio osmextract/osmextract.Rproj
```
Then in RStudio:

```{r, eval = FALSE}
file.edit("data-raw/geofabrik_zones.R")
```

Create a new script to document the code that generates the new object, e.g. for `bbbike`:

```{r, eval = FALSE}
file.edit("data-raw/bbbike_zones.R")
# or, even better, use
usethis::use_data_raw("bbbike_zones")
```

After you have created the new provider `_zones` file, it's time to add the provider to the package.

# Adding the new provider to the package

Once you have created your overview `_zones` file as outlined in the previous step, you need to modify the following files for the provider to be available for others:

- [data.R](https://github.com/ITSLeeds/osmextract/blob/master/R/data.R), where you'll need to document the new dataset;
- [globals.R](https://github.com/ITSLeeds/osmextract/blob/master/R/globals.R), where you'll need to add the new object name;
- [providers.R](https://github.com/ITSLeeds/osmextract/blob/master/R/providers.R), where you'll need to add the new object name in `oe_available_providers()` and `load_provider_data()`. 

# Documenting the provider

The final step is also the most fun: documenting and using the provider.
Add an example, mention it in the README and tell others about what this new provider can do!
If you want to ask for help on adding a new provider, feel free to open in a new issue in the [github repository](https://github.com/ITSLeeds/osmextract)! 

# Conclusion

This vignette talks through the main steps needed to extend `osmextract` by adding new OSM data providers.
To see the same information in code form, see the PR that implemented the `openstreetmap_fr` provider here: https://github.com/ITSLeeds/osmextract/commit/dbf131667a80e5a6837a6c8eb3b967075e1aba16
