---
title: "Add new OpenStreetMap providers"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Add new OpenStreetMap providers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(osmextractr)
```

As of summer 2020 there are 7 services providing bulk OSM datasets listed at [https://wiki.openstreetmap.org/wiki/Processed_data_providers](https://wiki.openstreetmap.org/wiki/Processed_data_providers).
This package is designed to make it easy to add new providers.
To add a new provider there are three main steps, outlined below.

# Adding a `provider_zones` object to the package

The first and hardest step is to create an R object analogous to the `test_zones` object shown below:

```{r}
names(test_zones)
test_zones[c(2, 8)]
```

The output shows the three most important column names in dataset:

1. The dataset name
2. The url endpoint where .pbf files associated with each zone can be downloaded
3. The geometry, representing the spatial extent of the dataset

The best way to start creating a new `_zones` object for a new provider is probably by looking at the code we wrote for the first and [test](https://github.com/ITSLeeds/osmextractr/commit/be3b48e7ed7ccd957e988bacad053161247b206d) providers in [`data-raw/geofabrik_zones.R`](https://github.com/ITSLeeds/osmextractr/blob/master/data-raw/geofabrik_zones.R)
The following commands will clone this repo and open the relevant file:

```bash
git clone git@github.com:itsleeds/osmextractr
rstudio osmextractr/osmextractr.Rproj
```

Then in RStudio:

```{r, eval=FALSE}
file.edit("data-raw/geofabrik_zones.R")
```

Create a new script to document the code that generates the new object, e.g. for `bbbike`:

```{r, eval=FALSE}
file.edit("data-raw/bbbike_zones.R")
```
