---
title: "Tips and tricks for using the package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tips and tricks for using the package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  fig.align = "center"
)

# save user's options and pars
user_options = options()
user_par = par(no.readonly = TRUE)

# save files in the tempdir
old_dd = Sys.getenv("OSMEXT_DOWNLOAD_DIRECTORY", tempdir())
Sys.setenv(OSMEXT_DOWNLOAD_DIRECTORY = tempdir())

its_pbf = file.path(
  osmextract::oe_download_directory(), 
  "test_its-example.osm.pbf"
)
file.copy(
  from = system.file("its-example.osm.pbf", package = "osmextract"), 
  to = its_pbf, 
  overwrite = TRUE
)

# set new options
options(width = 100)
```

This vignette presents a collection of useful tips and tricks we've gathered over the years for effectively using this package to read, download, and filter OpenStreetMap (OSM) extracts. First of all, let's load the relevant packages: 

```{r setup}
library(osmextract)
```

## How can I get OSM objects by node/way id number?

The example below demonstrates how to select a set of ways from an OSM extract, assuming you already know their OSM IDs:

```{r}
osm_id <- c("4419868", "6966733", "7989989", "15333726", "31705837")

out <- oe_get(
  place = "ITS Leeds",
  query = paste0(
    "SELECT * FROM lines WHERE osm_id IN (", paste0(osm_id, collapse = ","), ")"
  ), 
  quiet = TRUE
)
print(out, n = 0L)
```

## How can I convert the segments downloaded from OSM into a street network? 

Starting from version 0.7, the package includes a set of functions to automatically convert the output of `oe_get_network()` (and similar functions) into `sfnetwork` or `dodgr` objects. 

For example, the following command returns a `sfnetwork` object representing the walking network extracted from the toy ITS data included in the package: 

```{r}
sfn_walking = oe_get_sfnetwork(
  place = "ITS Leeds",
  mode = "walking", 
  quiet = TRUE
)
```

More precisely, the function runs the following operations: 

  1. It extracts the walkable streets from the toy ITS Leeds data included in the package; 
  2. It applies a series of preprocessing steps to standardise the values included in the `oneway` column and simplies the `highway` tag removing the `"_link"` suffix. See the `clean_output` argument in `oe_get_network` for more details. 
  3. It converts the data into `sfnetwork` class and applies two spatial morphers, namely `to_spatial_subdivision` and `to_spatial_smooth` to fix possible inconsistencies in the geometries and simplify the geometry structure. See also `?net_2_sfnet_undirected`; 
  4. If requested, converts the output into a directed network after duplicating bidirectional edges with reversed geometries. 
  
The following command runs similar operations (except for the spatial morphers) and returns a `dodgr_streenet` object: 

```{r}
dodgr_walking = oe_get_dodgrnetwork(
  place = "ITS Leeds",
  mode = "walking", 
  quiet = TRUE
)
```

See the help pages of the corresponding functions for more details. 

```{r, include=FALSE}
# reset par, options, and download directory
options(user_options)
par(user_par)
Sys.setenv(OSMEXT_DOWNLOAD_DIRECTORY = old_dd)
```
